\Title

### 2021-01-06 manuscript/manuscript.tex

\subsection{Source distributions}
\label{sec:source-distribution}

How to choose how many sources to use and where to locate them are not
trivial decisions that play an important role on the accuracy of the
predictions and the computational resources needed to fit the sources'
coefficients.

The chosen distribution of sources should simultaneously reproduce the measured
data on the survey points and be able to make accurate predictions of the
measured harmonic field on non observed locations.

A high number of sources homogeneously distributed on the survey region are
capable of reproducing the observed data. Nevertheless, too many sources will
make the Jacobian matrix too large, what could introduce computational problems
like the inability of storing the matrix on memory or a high computational time
needed to fit the sources' coefficients.
On the other hand, using very few sources will undoubtedly reduce the memory
needs and lower the computational time when fitting the sources' coefficients.
Nevertheless, it could be incapable of making accurate predictions or even
reproducing the measured data if the number of sources is too low.

The choice of the source distribution could also depend on the characteristic
of the survey and how the observation points are distributed on the region.
A ground survey usually consists in observation points located along irregular
paths plus some possible scattered points.
The survey region is usually heterogeneously covered, leaving areas without any
observation.
On the other hand, an airborne survey consists in observation points located
along almost straight flight lines. The measurements are usually taken at
a high frequency, so the observation points along the flight lines are very
close to each other.
This creates an anisotropy on the observation points: there are much more
observation points along the flight lines directions.

The most widely used source distributions are: a regular grid of point sources
and one point source beneath each observation point.

The \emph{regular grid} creates an homogeneous distribution of point sources
below the surveyed region.
The distance between sources is controlled by the grid spacing parameter.
Keeping the boundaries of the grid constant, reducing the spacing increases the
number of sources.

The \emph{source below data} layout locates one source below each
observation point, therefore the number of sources will be equal to the number
of data points.

When working with ground surveys, the \emph{regular grid} needs the grid
spacing to be sufficiently small so observed values can be accurately
reproduced by the source distribution.
This will create too many sources on areas where no observation has been
carried out.
In contrast, the \emph{source below data} layout is more likely to accurately
fit the observed data with much less sources, reducing the computational load
to fit the equivalent sources' coefficients.

When working with airborne surveys, the \emph{source below data} may
generate an anisotropic distribution of sources by locating too many sources
along the flight paths.
It could lead to unwanted aliases on the predicted values.
In contrast, the \emph{regular grid} won't generate aliases by homogeneously
locating sources.

We propose a new source distribution that could simultaneously reduce the
computational load and remove the drawbacks of the preceding ones: the
\emph{block averaged sources}.
It consist in:

\begin{enumerate}
    \item Dividing the region in blocks of equal size. The size of the blocks
        is controlled by a \emph{spacing} parameter.
        (see Fig.~\ref{fig:block-averaged-sources}(a)-(b)).
    \item Computing the median location of the observation points that fall
        inside each block. Blocks without any observation point are ignored.
    \item Locating one point source below each block averaged coordinate.
        (see Fig.~\ref{fig:block-averaged-sources}(c)).
\end{enumerate}

The number of sources created by this new distribution will be less or equal to
the ones created by the \emph{source below data} and the \emph{regular
grid} with the same spacing.
Moreover, the sources are only located beneath populated blocks, so sources
will be concentrated on regions where data has been gathered and no anisotropic
distribution of sources will be generated.

>>>>

\subsection{Location of sources}
\label{sec:source-distribution}

The ideal number of sources and their locations, both horizontally and
vertically, has been debated since the inception of the equivalent source
technique with \citet{dampney1969}.
The choices made regarding these parameters can play an important role on the
accuracy of the predictions and the computational resources needed to estimate
the coefficients.
An ideal distribution of sources should simultaneously be able to reproduce the
measured data on the survey points and also be able to make accurate
predictions on non-observed locations.

A large number of sources evenly distributed along the survey region are
capable of reproducing the observed data. Nevertheless, the computational load
can be prohibitive and such underdetermined problems are prone to overfiting
the data, leading to poor performance when making predictions. On the other
hand, using very few sources will undoubtedly reduce the computational
requirements but the equivalent source model could be incapable of reproducing
the measured data, leading to inaccurate predictions.

Particular survey characteristics also play a role in the choice of source
distribution.
In a ground survey, observation points are usually located along irregular
paths as well as some possible scattered points.
The coverage of the survey region is usually uneven, often leaving large areas
without any observation.
On the other hand, observations from airborne surveys are located along almost
straight flight lines.
Measurements are usually taken at
a high temporal frequency, leading to the observation points along the flight
lines that are several times closer than the flight line spacing.
This creates an anisotropy in the sampling, which can cause aliasing artifacts
in gridded products.

\subsubsection{Horizontal layout}

The most widely used layouts of equivalent sources are:

\begin{itemize}
  \item
    \emph{Regular grid}: a homogeneous distribution of point sources below the
    survey region at a constant depth. A padding region is often added to help
    reduce edge effect. In practice, often leads to underdetermined problems
    ($N<M$).
  \item
    \emph{Sources below data}: one equivalent source is placed below each
    observation point at a given depth (often dependent on the distance to
    neighboring data points). Therefore, the number of sources is be equal to
    the number of observations ($N=M$).
\end{itemize}

For ground surveys, the \emph{regular-grid} layout needs a sufficiently
small grid spacing to be able to fit the observed data.
This creates an unnecessarily large number of sources in areas where no
observations exist.
In contrast, the \emph{sources-below-data} layout is more likely to accurately
fit the observed data with many fewer sources, reducing the computational load.

For airborne surveys, the \emph{source-below-data} layout may
place an undesirably large number of sources along the flight paths.
This could lead to aliasing effects on the predicted values, such as the
stripes parallel to flight lines that are often observed when griding airborne
magnetic data.
In contrast, the \emph{regular-grid} layout can avoid this effect by evenly
distributing sources and using a continuous source layer (e.g.,
right-rectangular prisms or tesseroids).

We propose a new source layout that could simultaneously reduce the
computational load and remove some of the drawbacks of the regular-grid and
source-below-data layouts:

\begin{itemize}
  \item
    \emph{Block-averaged sources}: point sources are placed in the average
    position of data points that fall within specified spatial blocks
    (Fig.~\ref{fig:block-averaged-sources}). In
    short, this is done by:
    \begin{enumerate}
        \item Dividing the survey region in blocks of equal size
            (Fig.~\ref{fig:block-averaged-sources}a-b).
        \item Computing the median position of the observation points that fall
          inside each block. Blocks without any observation point are ignored.
        \item Assign one point source to the computed median position, shifted
          to a specified depth (Fig.~\ref{fig:block-averaged-sources}c).
    \end{enumerate}
\end{itemize}

The number of sources created by this new layout will be less than or equal to
(if the blocks contain a single data point)
the number of observations.
The overdetermined problems that arise from this layout have a lower
computational load and are less prone to overfiting since the complexity of the
model is lower.
Moreover, the block averaging process can balance the spacing between sources
along a flight line and between adjacent lines.
